---
tittle: rOceansNew - Examples
author: I. Montero-Serra
output:
  pdf_document: default
  html_document: default
---

## rOceans 
An R Package for exploring spatial trends in marine biodiversity and conservation. Here
have examples of functions to access and explore spatial patterns of marine biodiversity. 
The functions are separated in set #1 for single or few species analyses; and set #2 for
analyzing large data sets downloaded from OBIS or GBIF. 

Downloading rOceansNew from GitHub and loading the package
```{r}

devtools::install_github("monteroserra/rOceansNew")
library(rOceansNew)

````

# Functions 1: Accessing and exploring data for single / few marine species

The first function ()


```{r}



# Define target speices names: example with two corals
coral_species = c("Corallium rubrum", "Acropora palmata")

# Get occurrences 
corals = oceanDivDat(species_names = coral_species,
                     data_source = "OBIS_&_GBIF", 
                     remove_duplicates = T, 
                     remove_land_dots = T)

head(corals)

corals_presence = presencesRaster(occurrences=corals, 
                                         cell_size=5)

oceanMaps(corals_presence)


corals_richness = oceanRichness(occurrences=corals, 
                                cell_size=5, 
                                richness_map = T)


oceanMaps(corals_richness)


```

