---
tittle: rOceansNew - Examples
author: I. Montero-Serra
date: 22/Jun/2018
output: html_document
---

## rOceans 
An R Package for exploring spatial trends in marine biodiversity and conservation. Here
have examples of functions to access and explore spatial patterns of marine biodiversity. 
The functions are separated in set #1 for single or few species analyses and set #2 for
analyzing large data sets downloaded from OBIS or GBIF. 

*Currently, working on further set of functions to access and explore spatial data on climate change and other stressors as well as MPAs coverage. 

(1) Download rOceansNew from GitHub and load the package
```{r}

devtools::install_github("monteroserra/rOceansNew")
library(rOceansNew)

````

# Functions 1: Accessing and exploring data for single / few marine species

The first function ()


```{r}



# Define target speices names: example with two corals
coral_species = c("Corallium rubrum", "Acropora palmata")

# Get occurrences 
corals = oceanDivDat(species_names = coral_species,
                     data_source = "OBIS_&_GBIF", 
                     remove_duplicates = T, 
                     remove_land_dots = T)

head(corals)

corals_presence = presencesRaster(occurrences=corals, 
                                         cell_size=5)

oceanMaps(corals_presence)


corals_richness = oceanRichness(occurrences=corals, 
                                cell_size=5, 
                                richness_map = T)


oceanMaps(corals_richness)


```

