---
tittle: "Exploring ACropora corals abundance and diversity from GBIF and OBIS data"
author: "I. Montero-Serra, E. Aspillaga, K. Kaplan, V. Barbe & N. Barve"
date: 23/Jul/2018
rmarkdown::html_vignette:
  pdf_document:
    highlight: tango
    fig_caption: false
fontsize: 11pt
classoption: a4paper
vignette: >
  %\VignetteIndexEntry{Exploring patterns in Acropora corals abundance and diversity}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ">"
)
```

# Introduction

This document describes how to use rOceans, an R package for exploring spatial trends in marine biodiversity and conservation. Here we have a examples of usage for explore spatial patterns of biodiversity in scleractinian corals of the genus Acropora. 


(1) Download rOceansNew from GitHub and load the package

```{r}
invisible(devtools::install_github("monteroserra/rOceansNew"))
invisible(library(rOceansNew))

```

## Function #2 *oceanDataCheck* 
Allows merging, checking, filtering the raw data from OBIS and/or GBIF into a common standard database. It removes occurrences with inconsistent or inconmplete taxonomic information, land-based occurrences, and duplicate occurrences. 

For this set of functions, we first  need to access and download data from GBIF (www.gbif.org) and OBIS (www.iobis.org). 

Here I show an example with data of species within the genus *Acropora* that were directly downloaded from the website on June 2018. The data can be accessed from my GitHub repository "monteroserra/rOceanData" or using data()

```{r}
#Accesing raw datasets
data(Acropora_GBIF) 
data(Acropora_OBIS)

#Checking OBIS or GBIF individually
Acropora_OBIS_checked = oceanDataCheck(OBIS_occurrences = Acropora_OBIS, source = "OBIS")

Acropora_GBIF_checked = oceanDataCheck(GBIF_occurrences = Acropora_GBIF, source = "GBIF")
```

Merging OBIS and GBIF Datasets, and checking for taxonomic and geographic issues
```{r}
Acropora_Total_Checked = oceanDataCheck(GBIF_occurrences = Acropora_GBIF,
                                        OBIS_occurrences = Acropora_OBIS,
                                        source = "GBIF_&_OBIS")
```
