---
tittle: "Exploring sharks abundance and diversity from GBIF and OBIS data"
author: "I. Montero-Serra, E. Aspillaga, K. Kaplan, V. Barbe & N. Barve"
date: 23/Jul/2018
rmarkdown::html_vignette:
  pdf_document:
    highlight: tango
    fig_caption: false
fontsize: 11pt
classoption: a4paper
vignette: >
  %\VignetteIndexEntry{Exploring global patterns in sharks abundance and diversity}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ">"
)
```

\vspace{3em}


# Introduction

T this document describes how to use rOceansNew and R package for exploring spatial trends in marine biodiversity and conservation. Here
have examples of functions to access and explore spatial patterns of marine biodiversity. 

## Second example, exploring shark biodiversity trends of the genus Carharhinidae

#Compute a abundance grid

```{r}

data(Carcharhinidae_total_checked)

Carcharhinidae_abundance_10 = oceanAbundGrid(occurrences = Carcharhinidae_total_checked, cell_size=10)

Carcharhinidae_abundance_5 = oceanAbundGrid(occurrences = Carcharhinidae_total_checked, cell_size=5)

Carcharhinidae_abundance_1 = oceanAbundGrid(occurrences = Carcharhinidae_total_checked, cell_size=1)

Carcharhinidae_abundance_0.5 = oceanAbundGrid(occurrences = Carcharhinidae_total_checked, cell_size=0.5)
```

#Visualize abundance
```{r}
par(mfrow=c(2,2))
oceanMaps(Carcharhinidae_abundance_10, logScale=T, main="Carcharhinidae abundance (10x10)")
oceanMaps(Carcharhinidae_abundance_5, logScale=T, main="Carcharhinidae abundance (5x5)")
oceanMaps(Carcharhinidae_abundance_1, logScale=T, main="Carcharhinidae abundance (1x1)")
oceanMaps(Carcharhinidae_abundance_0.5, logScale=T, main="Carcharhinidae abundance (0.5x0.5)")
par(mfrow=c(1,1))
```

# Exploring envioronmental drivers of abundance patterns
```{r}
abundance_environ_df = invisible(oceanEnv(biod_grid = Carcharhinidae_abundance_1,
                                biodiv_metric = "abundance", 
                                plot=T, 
                                log_scale=T))
```

# Computing species abundance matrix - and several diversity metrics
```{r}
Carcharhinidae_diversity = oceanDiversity(occurrences = Carcharhinidae_total_checked, print=F)

# Visualizing diversity metrics
par(mfrow=c(2,2))
oceanMaps(Carcharhinidae_diversity[[2]], main = "Carcharhinidae richness")
oceanMaps(Carcharhinidae_diversity[[3]], main = "Carcharhinidae Shannon Diversity")
oceanMaps(Carcharhinidae_diversity[[4]], main = "Carcharhinidae Simpson Diversity")
par(mfrow=c(1,1))
```

# Exploring envioronmental drivers of abundance patterns
```{r}
diversity_env_df = invisible(oceanEnv(biod_grid = Carcharhinidae_diversity[[2]],
                                biodiv_metric = "richness", 
                                plot=T, 
                                log_scale=T))
```


Exploring biodiversity hotspots and vulnerability of these hotspots to warming
```{r}
Carcharhinidea_hotspots = oceanHotspots(biodiversity_grid = Carcharhinidae_diversity[[2]], 
                          hotspot_map=T, main = "Hotspots and coldspots of shark diversity")


hotspots_vulnerb <- oceanVulnerab(biodiversity_grid = Carcharhinidea_hotspots,
                     map_climtatic_impacts = F, plot_histograms = T)

```



